TRaM: Analyzing temporal framing in Intertemporal Choice

### 2015-12-15 paper/paperintro.tex
Write up intro and beginning of study 1

%\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}
\usepackage{booktabs,caption,fixltx2e}
%opening
\title{An exploration of how ratios affect time discounting}
\author{Daniel Wall}

@@

\begin{abstract}
	When selecting our shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
	This phenomenon -- known as intertemporal choice is the subject of this paper.
%	Choices involving time, intertemporal choices, are ubiquitous -- for instance choosing a shipping speed for your latest Amazon order. 
	People use ratios in everyday life -- intertemporal choices are no different, 
	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices. 
	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.

@@

All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible. 
Sometimes that itch may lead to us choosing to expedite our shipping at considerable cost. 
There are subtle ways in which the presentation of time can alter patience. 
For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
Marketers have used temporal frames in shipping scenarios.
For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
As per \citeA{Read2005} this frame will increase patience.

@@

Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster. 
It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.

Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

\section{Ratios and Intertemporal Choices}

@@

>>>>

%\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{apacite}
\usepackage{graphicx}
\usepackage{booktabs,caption,fixltx2e}
%opening
\title{TRaM: Analyzing temporal framing in Intertemporal Choice}
\author{Daniel Wall}

@@

\begin{abstract}
	When selecting a shipping speed on Amazon we are faced with a choice between getting our package sooner but paying more money, or getting our package later and paying less money. 
	The general phenomenon underlying our expedited shipping is impatience, we want good things immediately and we must be compensated for waiting. 
	Previous research on intertemporal framing has shown that displaying times as dates -- January 23rd -- as opposed to delays -- 6 weeks -- increases patience. 
	Other research has shown that people use both differences and ratios of time in intertemporal choices. 
%	These two results suggest that framing times as ratios alters patience. 
%	In three experiments I manipulate the temporal presentation from the standard -- 1 day versus 5 days --  to making the ratio explicit -- 1 day or 5 times as long.  
	Across three experiments I show that making temporal ratios explict leads to differing levels to increase impatience in standard  intertemporal choices framed as gains, intertemporal choices framed as losses, and for an explicit shipping scenario. 
	Taken together these results suggest that framing delays as ratios leads to greater impatience.

%
%	People use ratios in everyday life -- intertemporal choices are no different, 
%	recent cognitive models of intertemporal choice show that people calculate ratios of both money and time. 
%	These models account for diminishing impatience -- when items are moved into the future, people are more patient. 
%	Our goal is twofold: 1) to show that previous models of intertemporal choice do not properly account for temporal ratios and present a conceptual model of time discounting, and  2) by explicitly manipulating how people use ratios of time, we can alter how people make intertemporal choices. 
%	This leads to novel predictions for how to optimally frame intertemporal choices to induce either patience or impatience. 
%	Further we show that in a hypothetical shipping scenario that when temporal ratios are explicit, people are more impatient.

@@

All of us have felt the itch to get the new purchase we made on Amazon as quickly as possible; sometimes leading us choosing expedited shipping, even at considerable cost. 
Generally on Amazon we see shipping times as the number of days until you'll receive your package, e.g.  Receive your package in 2 days. 
However, an identical choice to the one above would be Recieve your package on December 16th. 
\citeA{Read2005} showed that people are more impatient when times are framed as days as compared framed as to dates. 
Accordingly when shopping online delays are occasionally framed as dates.

However a firm may want consumers to be impatient and choose the expedited shipping option. 
If a firm's inventory is high they may want to nudge consumers to ship their items faster. 
It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds.

How the monetary and temporal aspects of an intertemporal choice are framed lead to varying degrees of patience.
Further, recent models of intertemporal choice have posited that people use ratios in intertemporal choices \cite{Read2013, MarzilliEricson2015}. 
%For instance \citeA{Read2013} found that by making ratios in the 
However, to the best of my knowledge, no study has investigated how explicitly framing time as a ratio can alter patience.
Across 3 studies we show that by making ratios explicit people are more impatient in intertemporal choices. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios.
This result holds for intertemporal choices framed as gains, losses, and in an explict shipping scenario. 
%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

%
%There are subtle ways in which the presentation of time can alter patience. 
%For instance people are more impatient when delays are calendar dates as opposed to delays into the future \cite{Read2005}. 
%Marketers have used temporal frames in shipping scenarios.
%For example as opposed to stating that you will receive your package in 2 weeks, Amazon sometimes states that you receive your package on January 23rd. 
%As per \citeA{Read2005} this frame will increase patience.

@@

%Firms could, however, prefer that a consumer is impatient when choosing a shipping speed. 
%If a firm's inventory is low on a particular item they may not profit from the extra cost that it takes to expedite shipping, however if their inventory is high they may want to nudge consumers to ship their items faster. 
%It is important to understand what variables consumers are using and how they are making intertemporal choices to create robust and adaptive choice environments. 
%This understanding how to optimally display information leads to novel predictions for how to frame shipping speeds. 
%
%Recent models of intertemporal choice have posited that people use ratios in intertemporal choices [ITCH, DRIFT]. 
%Across [] studies we show that how people make intertemporal choices is dependent on how they perceive ratios. 
%Further we show that by making the ratios between times explicit we can induce a higher reliance on ratios. 
%If the temporal ratio is explicit people are more impatient than when the ratio is implicit.

\section{Intertemporal Choices}

Intertemporal choices in marketing and psychology are generally studied by choices between a smaller sooner and larger later amounts of money \cite{Zauberman2014}.

There are multiple categories of research on these intertemporal choices (see \citeA{Read2013} for an overview) however the current paper belongs to research which investigates how impatience is altered by option frames. 
An example of option framing in intertemporal choice is the hidden zero effect -- when the implicit zeros in intertemporal choices are made explicit people are more patient \cite{Magen2008}. 
For instance when a choice between ``receiving \$100 now'' and ``\$110 in 6 months'' is reframed to be  ``receiving \$100 now \textit{and nothing in six months}'' and ``\$110 in 6 months \textit{and nothing now}'' people are more patient.

Additionally \citeA{Read2013} show that making the ratio between amounts salient affects intertemporal choices. 
Specifically when the magnitude of amounts is small, reframing the amounts as interest rates leads to increased patience. Taking the above example, when reframed as an interest rate, e.g.  a choice between  ``receive \$100 now'' and ``receive 10\% more in six months'' people are more patient. \citeA{Read2013} posit that intrest framed amounts lead to an increased reliance on ratios.

Other work by \citeA{Read2005} shows that how a times are framed in intertemporal choices can alter patience. 
Specifically they find that by reframing the delay above to be a date induces greater patience. 
Taking the above example, people are more patient when reframed it as a choice between  ``receive \$100 now'' and ``receive \$110 on June 15''.
This demonstrates that changing how times are framed can affect patience.

In other work on intertemporal choices, \citeA{MarzilliEricson2015} show that in  standard framed intertemporal choices, people rely on both differences and ratios between both monetary amounts and times.
 Their model -- Intertemporal Choice Heuristic (ITCH) -- does better at predicting out of sample intertemporal choices than classic models of intertemporal choices such as hyperbolic or exponential disounting. 
As seen in Equation~\ref{eq:itch}  the functional form of the itch model with estimated parameters shows that an increased reliance on ratios in time leads to steeper discounting.

[What are they?]

This paper combines the work of \citeA{Read2013} -- which shows that framing amounts as ratios alters patience -- and \citeA{MarzilliEricson2015} -- which shows that people rely on differences and ratios in standard intertemporal frame as well as the work  of  \citeA{Read2005} which shows that how times are framed can alter patience. 
Specifically, I focus on how framing times as ratios can decrease patience.
When delays are sufficiently small, and temporal ratios are made explicit, our attention will be focused on how much \textit{longer} we will wait. 
This focus on how much longer we will wait will lead to decreased patience when times are framed as ratios.

I operationalize this prediction across three studies.
Study 1 investigates how making temporal ratios explicit in intertemporal choices leads to greater impatience.

Study 2 extends the results of Study 1 to losses. 
Although most intertemporal choice studies use gains, a large number of real world intertemporal choices involve losses \cite{Hardisty2013}. 
For instance deciding to pay off your credit card incurs greater losses if you wait -- the compounding interest -- than if pay it off immediately.
Intertemporal choices framed as losses take the form of choices between ``Pay \$100 immediately'' and  ``Pay \$110 in 6 months'' \cite{Hardisty2015}.
Further most studies investigating framing effects on intertemporal choices are restricted to the domain of gains. 
Study 2 investigates how framing times as ratios leads to decreased patience in losses.

Study 3 extends the results from Study 1 and Study 2 to intertemporal choices explicitly framed as shipping scenarios.
Shipping scenarios are displayed as a choice between ``Receive your items in 2 days and pay \$8'' and ``Receive your items in 10 days and pay \$4''
Although choosing a shipping speed is similar to the intertemporal choices there is a subtle, yet possibly significant difference between the two.
Shipping decisions are paying more now to receive an item sooner whereas the intertemporal choices shown in study 2 are paying something smaller now to avoid paying something larger later.

\subsection{Hypothesis}

\subsubsection{Hypothesis 1}
According to the ITCH model people rely on ratios and differences in temporal choices. 
A prediction of the model is that the higher the weight given to the temporal ratio the more impatient participants will be.

\textit{Hypothesis 1} In gain intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.

\subsubsection{Hypothesis 2}
Similar to hypothesis 1 when framed as losses making temporal ratios explicit people will be more impatient.

\textit{Hypothesis 2} In loss intertemporal choices, when times are expressed as ratios, people will be more impatient than when expressed as delays.

\subsubsection{Hypothesis 3}
Similar to hypothesis 2 when framed as a shipping scenario making temporal ratios explicit people will be more impatient.

\textit{Hypothesis 3} In shipping scenarios, when times are expressed as ratios, people will be more impatient than when expressed as delays.

\section{Study 1}
In study 1 participants chose between smaller sooner and larger later options. 
These options were displayed in one of two frames: implicit temporal ratios and explicit temporal ratios. 
Study 1 tests Hypothesis 1 above

\subsection{Method}
Two hundred participants from Amazon Mechanical turk we recruited to take part in a ``Decision Making Survey''. 
Although the validity of Mechanical Turk has been called into question, \citeA{Paolacci2010} finds that classic effects in decision making replicate. 
Furhter, participants who have less than 50 HITs accepted or  who have less than 95\% of their HITs accepted were not allowed to participate in the study.

Participants completed informed consent, then were given a test to ensure they understood the paradigm.
They then were assigned to one of the two conditions: implicit ratio or explicit ratio. 
They then were shown the intertemporal choices in Table []. 
Each intertemporal choice was shown in the following format.

Insert a qualtrics picture here.

Choices were displayed in a random order to each participant.

After completeing their choices participants completed demographic information.

The dependent variable in this experiment was the number of times a participant chose the larger later -- e.g. more patient -- reward.

each participant chose the larger-later reward ?LL?.As Table 2 shows, there was a strong date/delay effect: Twice as many people chose LL in the Date condition than in the Week or Month conditions. Differences across conditions were tested formally with Kruskal- Wallis tests. An overall test confirmed a signifi- cant difference between the conditions (?2?2? = 21?1,

@@

%\begin{equation}\label{eq:itch}
%	P(LL) = L \left(\beta_I + \beta_{xA}(x_2 - x_1) + \beta_{xR} \frac{x_2 - x_1}{x^*} + \beta_{tA}(t_2 - t_1) + \beta_{tR} \frac{t_2 - t_1}{t^*}\right)
%\end{equation}
